Evaluation of +S2 sea ice forecast and some other better title

### 2025-02-24 analysis/paper/access-ice.tex
udaptes error metrics

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\IfFileExists{pdfcomment.sty}
{
  \usepackage{pdfcomment}
  \newcommand{\criticmarkupadd}[1]{\pdfmarkupcomment[markup=Highlight,color={0.838431  0.913725  0.734902}]{##1}{}}
  \newcommand{\criticmarkuprm}[1]{\pdfmarkupcomment[markup=Highlight,color={0.887059  0.673725  0.673725}]{##1}{}}
  \newcommand{\criticmarkupmark}[1]{\pdfmarkupcomment[markup=Highlight,color={1  0.99216  0.21961}]{##1}{}}
  \newcommand{\criticmarkupcomm}[1]{\pdfcomment[icon=Comment, color={0.89804  0.69020  0}]{##1}}
}{

  \newcommand{\criticmarkupadd}[1]{\{++{##1}++\}}
  \newcommand{\criticmarkuprm}[1]{\{-{}-{##1}-{}-\}}
  \newcommand{\criticmarkupmark}[1]{\{=={##1}==\}}
  \newcommand{\criticmarkupcomm}[1]{\{>{}>{##1}<{}<\}}
}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Untitled},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Untitled}
\author{}
\date{}

\begin{document}
\maketitle
\begin{abstract}
balbalbal
\end{abstract}

\section{Introduction}\label{introduction}

\criticmarkupmark{\textbf{Antarctic sea ice general description}.
Seasonality, trends.}

\criticmarkupmark{\textbf{Importance}. Impacts on life. Potential
impacts on weather. Impacts on climate. Impacts on ocean
circulation.(Justify why it's important to study and understand)}

\criticmarkupmark{\textbf{Prediction systems.} Why is it important:
operations \criticmarkupcomm{Discuss with Phil}; science needs (we need
good models). What is available. Discussion on S1 \criticmarkupcomm{Can
we discuss previous S1 results? Laura's paper is not public.}}

\criticmarkupmark{\textbf{Objective.} Evaluate S2.}

\section{Data and methods}\label{data-and-methods}

All datasets are regrided using bilinear interpolation to a common grid,
which is a polar sterographic grid with approximatedly equal area of
25km².

\subsection{Description of the model}\label{description-of-the-model}

ACCESS-S2 is the next saeasonal prediction system after ACCESS-S1. It
uses virtually the same exact model configuration except for some model
versions being different. The main difference are the initial
conditions. While S1 used initial conditions from the UK Met Office, S2
uses the Bureau's own system. \criticmarkupmark{ Description of the
system }\criticmarkupcomm{Atmosphere and ocean. Highlight no ice data
assimilation. Highligh that the hope is that ice will respond to
SST/atmosphere DA}

For evaluation we use hindcst for the period
1981--2023\criticmarkupcomm{Check}. Anomalies will be taken with respect
to the 1981--2011 climatology computed from the reanalysis. Climatology
is smoothed with a 11 day running mean.

\subsection{Verification datasets}\label{verification-datasets}

There is not a lot of data on sea ice properties, especially for things
like thickness, age, etc. However there are relatively reliable
satellite-derived estimates of sea ice concentration, which estimates
the proportion of each grid area that is covered with ice. These
products are not perfect and there still exists a fair bit of
observation al uncertainty. To account for this uncertainty we use
multiple datasets.

\subsubsection{Bootstrap}\label{bootstrap}

\criticmarkupcomm{From https://nsidc.org/data/nsidc-0079/versions/4}
This sea ice concentration data set was derived using measurements from
the Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7
satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors
on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and
-F13 satellites. Measurements from the Special Sensor Microwave
Imager/Sounder (SSMIS) aboard DMSP-F17 are also included. The data set
has been generated using the Advanced Microwave Scanning Radiometer -
Earth Observing System (AMSR-E) Bootstrap Algorithm with daily varying
tie-points. Daily (every other day prior to July 1987) and monthly data
are available for both the north and south polar regions. Data are
gridded on the SSM/I polar stereographic grid (25 x 25 km) and provided
in two-byte integer format. Data coverage began on 01 November 1978 and
is ongoing through the most current processing, with updated data
processed several times annually.

(Comiso 2023)

\subsubsection{NASA Team}\label{nasa-team}

\criticmarkupmark{missing}

\subsubsection{CDR}\label{cdr}

NOAA/NSIDC's Climate Data Record V4 combines the Bootstrap and NASA Team
estimates in an attempt to overcome each algorithm's weaknesses.
\criticmarkupcomm{More detail on how are they combined and their
properties.} As such, it is is not a truly independent dataset.

Meier, W. N., F. Fetterer, A. K. Windnagel, and S. Stewart. 2021.
NOAA/NSIDC Climate Data Record of Passive Microwave Sea Ice
Concentration, Version 4. {[}Indicate subset used{]}. Boulder, Colorado
USA. NSIDC: National Snow and Ice Data Center
https://doi.org/10.7265/efmz-2t65. {[}Date Accessed{]}
\criticmarkupcomm{Add to zotero}

\subsubsection{ERA5}\label{era5}

\criticmarkupmark{From the CDS websiste:} ERA5 is the fifth generation
ECMWF reanalysis for the global climate and weather for the past 8
decades. Data is available from 1940 onwards. ERA5 replaces the
ERA-Interim reanalysis.

This parameter is the fraction of a grid box which is covered by sea
ice. Sea ice can only occur in a grid box which includes ocean or inland
water according to the land-sea mask and lake cover, at the resolution
being used. This parameter can be known as sea-ice (area) fraction,
sea-ice concentration and more generally as sea-ice cover. In ERA5,
sea-ice cover is given by two external providers. Before 1979 the
HadISST2 dataset is used. From 1979 to August 2007 the OSI SAF (409a)
dataset is used and from September 2007 the OSI SAF oper dataset is
used. Sea ice is frozen sea water which floats on the surface of the
ocean. Sea ice does not include ice which forms on land such as
glaciers, icebergs and ice-sheets. It also excludes ice shelves which
are anchored on land, but protrude out over the surface of the ocean.
These phenomena are not modelled by the IFS. Long-term monitoring of sea
ice is important for understanding climate change. Sea ice also affects
shipping routes through the polar regions.

\criticmarkupcomm{Why is ERA5 better or worse than
satellites?\textless{} \textless\}

\subsection{Error measures}\label{error-measures}

@@

\subsubsection{Sea ice extent}\label{sea-ice-extent}

Sea Ice Extent is defined as the area of ocean covered with at least
15\% ice. This threshold is motivated by the limitations in satellite
retrieval, which is increasingly unreliable for low sea ice conditions.

Sea Ice Extent is a rough global measure, but a model could have
relatively accurate extent of ice but with different distributions. We
use two other measures to account for these errors.

We compute Root Mean Square Error of sea ice concentration anomalies.

We also compute the Integrated Ice Edge Error (IIEE) (Goessling et al.
2016). This is defined as the area in which the model misspredicts sea
ice concentration being above or belog 15\% ice. That is, dichotomise
sea ice concentration into areas with more and less than 15\% sea ice
both in the forecast and observations; the IIEE is the area in which
forecast and observations differ.

\section{Results and discussion}\label{results-and-discussion}

\subsection{Reanalysis}\label{reanalysis}

\subsubsection{Bias}\label{bias}

@@

\centering{

\includegraphics{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}

@@

\caption{\label{fig-hindcast-extent}Median sea ice extent for al
hindcasts initialised the first of the month for ACCESS-S2 and ACCESS-S1
in colours representing the start month. In black, the}

\end{figure}%

Figure~\ref{fig-hindcast-extent} a shows the seasonal cycle of Sea Ice
Extent for the ACCESS-S2 hindcast and 2 observational datasets.
ACCESS-S2 shows a severe low extent bias, especially in the late
summer-early autumn. This is due primarily to a faster and longer melt
season between January and March and slower growth during March and
April This is then balanced with faster growth between May and July
(Fig.~\ref{fig-mean-growth}). This bias is common in climate models
{[}ref!{]} and it's been linked to xxxx??

Comparing ACCESS-S2 with ACCESS-S1, the latter has a smaller bias,
especially at low lags (Fig.~\ref{fig-hindcast-extent}) even though the
typical growth rates are very similar between both models
(Fig.~\ref{fig-mean-growth}). At larger lags, ACCESS-S1's bias in summer
and autumn is very similar to ACCESS-S2`s. This suggests that this lower
sea ice state is closer to the models' equilibrium, indicating that it
is an issue with model formulation that was being corrected by the data
assimilation system in ACCESS-S1.

At long lags, sea ice extent loses most of the initial condition memory
and reverts to the model's preferred equilibrium state. Therefore we can
estimate the latter using the hindcasts with the largest possible lag,
which is shown in triangles in Figure~\ref{fig-hindcast-extent} for the
same dates as the initial conditions. The difference between the two is
the effect of the data assimilation.

The equilibrium of ACCESS-S1 and ACCESS-S2 is very similar (comparing
the triangles in each panel in Fig.~\ref{fig-hindcast-extent}), owing to
both having the same model formulation. From June to October, in
ACCESS-S2 circles move away from triangles and towards observations,
indicating that the information from the ocean and atmosphere data
assimilation is getting to sea ice and affecting the initial conditions.
The rest of the year, there is little if any difference between circles
and triangles in ACCESS-S2, indicating that almost no data assimilation
is taking place and the sea ice component of the model is virtually
free-running.

@@

\centering{

\includegraphics{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}

@@

\end{figure}%

@@

\centering{

\includegraphics{access-ice_files/figure-pdf/fig-bias-1.pdf}

@@

\caption{\label{fig-bias}ACCESS-S2 reanalysis sea ice concentration bias
compared with NSIDC sea ice concentration.}

\end{figure}%

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[tc, manuscript]{copernicus}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage[inline]{trackchanges}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Evaluation of +S2 sea ice forecast and some other better title},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={red},
  pdfcreator={LaTeX via pandoc}}

\title{Evaluation of +S2 sea ice forecast and some other better title}

\Author[1,2]{Elio}{Campitelli}
\Author[1,2]{Ariaan}{Purich}
\Author[1,2]{Julie}{Arblaster}
\Author[3]{Eun-Pa}{Lim}
\Author[3]{Matthew}{Wheeler}
\Author[3]{Phillip}{Reid}

\affil[1]{School of Earth, Atmosphere and Environment, Monash
University, Australia}
\affil[2]{Securing Antarctica's Environmental Future, Monash University,
Australia}
\affil[3]{Bureau of Meteorology}

\runningtitle{+S2 sea ice forecast}

\runningauthor{Campitelli et al.}

\correspondence{Elio\ Campitelli\ (elio.campitelli@monash.edu)}

\begin{document}

\received{}
\pubdiscuss{} %% only important for two-stage journals
\revised{}
\accepted{}
\published{}

%% These dates will be inserted by Copernicus Publications during the typesetting process.

\firstpage{1}
\maketitle
\begin{abstract}
The abstract goes here. It can also be on \emph{multiple lines}.
\end{abstract}
\copyrightstatement{The author's copyright for this publication is
transferred to institution/company.}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[enhanced, boxrule=0pt, frame hidden, sharp corners, interior hidden, borderline west={3pt}{0pt}{shadecolor}, breakable]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Unlike Arctic sea ice, which has been steadily retreating at least since
the start of satellite records in the early 80s, Antarctic sea ice has
experienced a slightly increasing trend that models systematically
failed to reproduce and puzzled researchers \citep{hobbs2016}. Then, in
2016 Antarctic sea ice extent dropped precipitously and has been at low
and record low levels since, highlighting even more our lack of
understanding of Antarctic sea ice variability and change
\citep{espinosa2024}.

Antarctic sea ice is in a remote location but the potential impacts of
its variability extends far into the lower latitudes, affecting the
circulation of both the ocean and the atmosphere, ocean carbon uptake
and biological processes. It's likely that a reduction in sea ice extent
would reduce the meridional temperature gradient which in turn would
reduce the strength and location of the eddy-driven jet (and vice versa
for an increase in sea ice), which is a crucial component of weather
events. In the northern hemisphere, it has been hypothesised that the
reduction in Arctic sea ice has lead to a weaker, wavier jet, increasing
the frequency of extreme events \citep{barnes2015}. There is some
evidence of this effect in the Southern Hemisphere. Models show a
weakened vortex in the middle atmosphere, increased vertical wave flux
and greater zonal wave 1 amplitude in Springs following a low sea ice
maximum. But models don't fully agree on the exact phase of the zonal
wave response and observations don't show a clear signal
\citep{rea2024}. The jet response to sea ice loss is also very model
dependent, particularly sensitive to its the climatological location and
relatively small compared with the direct effect of CO2 increase and Sea
Surface Temperature (SST) warming \citep{ayres2019}. On climate scales,
it is thought that the temperature signal of a reduction in Antarctic
sea ice mirrors a ``mini global warming'' and experiments also predict a
weakening and northerly shift of the eddy-driven
\citep{raphael2011, england2018, england2020, ayres2022}.

Correctly modelling Antarctic sea is not only necessary for process
understanding and climate projections to inform adaptation strategies.
Accurate seasonal to subseasonal forecast are crucial for operations in
and around the Antarctic continent such as scientific missions,
fisheries and tourism \citep[ \citet{wagner2020}]{desilva2020}. However,
producing accurate Antarctic sea ice forecasts has been challenging due
to model biases and inherent large variability and complexity and it has
lagged behind Arctic sea ice forecasts \citep{zampieri2019}. Dynamical
seasonal forecasts of summer Antarctic sea ice have been shown to
perform worse than relatively simple statistical methods
\citep{massonnet2023}, which also underscores the need for better
understanding of sea ice dynamics.

In this work we evaluate sea ice forecasts produced by ACCESS-S2, the
Australian Bureau of Meteorology subseasonal-to-seasonal prediction
system. This evaluation will inform future work with the model as a
reseach tool and explore the potential of using it as a sea ice
forecasting tool.

\hypertarget{data-and-methods}{%
\section{Data and methods}\label{data-and-methods}}

\hypertarget{s2}{%
\subsection{ACCESS-S2}\label{s2}}

ACCESS-S2 \citep{wedd2022} is the Australian Bureau of Meteorology
current subseasonal-to-seasonal prediction system which replaced
ACCESS-S1 \citep{hudson2017} in October 2021. Both ACCESS-S2 and
ACCESS-S1 consist on the Global Atmosphere 6.0 (GA6)
\citep{williams2015, waters2017}, Global Land 6.0
\citep{best2011, waters2017}, Global Ocean 5.0
\citep{madec2013, megann2014} and Global Sea Ice 6.0 {[}CICE;
\citet{rae2015}{]} model components. The atmosphere has a N216
horizontal resolution (\textasciitilde60km in the mid-latitudes) with 85
levels. The land model uses the same horizontal grid with 4 soil levels.
The ocean component has a 1/4º resolution with 75 vertical levels. The
ice component --based on CICE version 4.1-- has the same resolution than
the ocean and 5 sea ice categories as well as an open water category.

Both systems take atmospheric initial conditions derived from
ERA-interim \citep{dee2011} for the hindcast and from the Bureau's
operational analysis for real-time forecasts. The main difference
between the two are the ocean and ice initial conditions. ACCESS-S1
initial conditions come from the Met Office FOAM system, which uses a
multivariate, incremental three-dimensional variational (3D-Var),
first-guess-at-appropriate-time (FGAT) data assimilation scheme
\citep{waters2015} and assimilates sea surface temperature (SST), sea
surface height (SSH), in situ temperature and salinity profiles, and sea
ice concentration. ACCESS-S2, instead, runs from initial conditions
generated by the in-house data assimilation scheme described in
\citet{wedd2022}. This scheme is a weakly coupled ensemble optimal
interpolation method and assimilates temperature and salinity profiles
from EN4 \citep{good2013} for the hindcast and from the WMO Global
Telecommunication System (GTS) and both the Coriolis and USGODAE Global
Data Assembly Centers (GDACs) for the real-time forecast. SSTs are
nudged to Reynolds OISSTv2.1 \citep{reynolds2007} for the hindcast and
to the Global Australian Multi-Sensor SST Analysis (GAMSSA; Zhong and
Beggs 2008) for the real-time forecast in areas where SSTs are over 0ºC.
Relevant for this work, sea ice concentrations are not assimilated.

The ACCESS-S2 hindcast used in this study runs for the period
1981--2018. Each forecast consists of 9 ensemble members, 3 of which run
for 279 days and 6 for 42 days. Anomalies are be taken with respect to
the 1981--2011 climatology computed from the daily ACCESS-S2 reanalysis
and smoothed with a 11 day running mean.

The ACCESS-S1 hindcast runs for the period 1990--2012, with 7-member
forecasts integrated for 217 days. No reanalysis is available so we are
not able to compute a climatology for this forecast, therefore we don't
show metrics based on anomalies.

\hypertarget{verification-datasets}{%
\subsection{Verification datasets}\label{verification-datasets}}

For verification we use satellite-derived estimates of sea ice
concentration, which estimates the proportion of each grid area that is
covered with ice. Different datasets derived using different algorithms
provide different estimates, each with its own biases and uncertainties.
\citet{meier2019} estimated that the inter-product uncertainty of sea
ice extent is of the order of 1 million km\textsuperscript{2}. As will
be shown below, this spread is minimal compared with the errors in the
ACCESS-S2 and ACCESS-S1 forecasts, so the conclusions are independent of
the dataset used.

We use NOAA/NSIDC's Climate Data Record V4 {[}CDR; \citet{meier2014}{]}
as the primary verification dataset. It takes the maximum value of the
NASA Team \citep{cavalieri1984} and NASA Bootstrap \citep{comiso2023}
sea ice concentration products to reduce their low concentration bias
\citep{meier2014, meier2021}. Both source algorithms use data from the
Scanning Multichannel Microwave Radiometer (SMMR) on the Nimbus-7
satellite and from the Special Sensor Microwave/Imager (SSM/I) sensors
on the Defense Meteorological Satellite Program's (DMSP) -F8, -F11, and
-F13 satellites. The data has a resolution of 25 by 25km and daily from
1978 onwards.

The European Organisation for the Exploitation of Meteorological
Satellites (EUMETSAT) Ocean and Sea Ice Satellite Application Facility
(OSI SAF) \citep{OSISAF} is another satellite-derived sea ice
concentration product. It is based on mostly the same sensors as the NOA
CDR but computed independently using different algorithms. Figures
prepared with this dataset are provided in the supplementary material.

\hypertarget{error-measures}{%
\subsection{Error measures}\label{error-measures}}

@@

Sea Ice Extent is defined as the area of ocean covered with at least
15\% ice. This threshold is motivated by the limitations in satellite
retrieval, which is increasingly unreliable for lower sea ice
conditions.

Pan-Antarctic Sea Ice Extent serves as a rough global measure of the
amount of sea ice but it doesn't take into account the spatial
distribution. A model could have relatively accurate extent of ice but
with different distributions. To account for location errors, we
computed the Root Mean Squared Error (RMSE) of sea ice concentration
anomalies and the Integrated Ice Edge Error {[}IIEE;
\citet{goessling2016}{]}.

We compute RMSE as the square root of the average squared differences
between forecasted and observed sea ice concentration anomalies. We
compute a pan-Antarctic RMSE by averaging over the whole domain, and
also a zonally-varying RMSE by averaging over 24 15° longitude slices
around Antarctica.

The IIEE is defined as the area in which the model misspredicts sea ice
concentration being above or belong 15\% ice. That is, dichotomise sea
ice concentration into areas with more and less than 15\% sea ice both
in the forecast and observations; the IIEE is the area in which forecast
and observations differ.

All error measures were computed on the NOAA/NSIDC CDRV4 domain grid and
projection to which model output was bilinearly interpolated.

\hypertarget{results-and-discussion}{%
\section{Results and discussion}\label{results-and-discussion}}

\hypertarget{reanalysis}{%
\subsection{Reanalysis}\label{reanalysis}}

\hypertarget{bias}{%
\subsubsection{Bias}\label{bias}}

@@

{\centering \includegraphics{access-ice_files/figure-pdf/fig-hindcast-extent-1.pdf}

@@

\caption{\label{fig-hindcast-extent}Median sea ice extent for al
hindcasts initialised the first of the month for ACCESS-S2 and ACCESS-S1
in colours representing the start month. In black, the median sea ice
extent of NOAA/NSIDC CDRV4.}

\end{figure}

Figure~\ref{fig-hindcast-extent} a shows the seasonal cycle of Sea Ice
Extent for the ACCESS-S2 hindcast and NOAA/NSIDC CDRV4 ACCESS-S2 shows a
severe low extent bias, especially in the late summer-early autumn. This
is due primarily to faster melt during a longer melt season between
January and March and slower growth during March and April. This is then
balanced with faster growth between May and July
(Fig.~\ref{fig-mean-growth}). Many sea ice models exhibit this
systematic underestimation during the sea ice minimum and early freezing
season \citep{massonnet2023} and could indicate problems in the
representation of thermodynamics in the model \citep{zampieri2019}.

ACCESS-S1 has a smaller low bias than ACCESS-S2, especially at low lags
(Fig.~\ref{fig-hindcast-extent}) even though the typical growth rates
are very similar between both models (Fig.~\ref{fig-mean-growth}). At
larger lags, ACCESS-S1's bias in summer and autumn is very similar to
ACCESS-S2's. This suggests that this lower sea ice state is closer to
the models' equilibrium, indicating that it is an issue with model
formulation that was being corrected by the data assimilation system in
ACCESS-S1 and not in ACCESS-S2.

At large lags, sea ice extent loses most of the initial condition memory
and reverts to the model's preferred equilibrium state. Therefore we can
estimate the latter using the hindcasts with the largest possible lag,
which is shown in triangles in Figure~\ref{fig-hindcast-extent} for the
same dates as the initial conditions. The difference between the two is
the effect of the data assimilation.

The equilibrium of ACCESS-S1 and ACCESS-S2 is very similar (comparing
the triangles in each panel in Fig.~\ref{fig-hindcast-extent}), owing to
them having the same model formulation. From June to October, in
ACCESS-S2 circles are closer to observations than to the triangles,
indicating that the information from the ocean and atmosphere data
assimilation is getting to sea ice and improving the initial conditions.
The rest of the year, there is little if any difference between circles
and triangles in ACCESS-S2, indicating that almost no data assimilation
is taking place and the sea ice component of the model is virtually
free-running. In ACCESS-S1, on the other hand, circles are very close to
observations year-round.

@@

{\centering \includegraphics{access-ice_files/figure-pdf/fig-mean-growth-1.pdf}

@@

\end{figure}

@@

{\centering \includegraphics{access-ice_files/figure-pdf/fig-bias-1.pdf}

@@

\caption{\label{fig-bias}Mean sea ice concentrations 1-month lag monthly
mean ACCESS-S2 forecast bias compared with NSIDC.}

\end{figure}

Figure~\ref{fig-bias} shows the difference of monthly mean sea ice
concentrations between NOAA/NSIDC CDRV4 and ACCESS-S2 hindcasts at the
shortest lag. From October to May, the model underestimates sea ice
concentrations pretty much everywhere there is ice except for the deep
Weddell Sea in April and May, where sea ice concentrations saturate to 1
both in the observations and forecasts. In winter, the differences are
mostly on the sea ice edge, with light positive bias in the African
sector of East Antarctica and negative bias around the Indian Ocean
sector which partially compensate, resulting in the near-zero extent
bias seen in those months (Fig.~\ref{fig-hindcast-extent}).

\hypertarget{anomalies}{%
\subsubsection{Anomalies}\label{anomalies}}
